---
title: "random notes"
author: "Kaitlyn Westra"
date: "11/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r read-data}
M2_all <- read.delim('./M2_all.txt')
M_tmn_wide_st <- read.delim('./M_tmn_wide_st.txt')
glimpse(M2_all)
glimpse(M_tmn_wide_st)
#M_tmn_long <- read.delim('./M_wide_all.txt.zip')
```



Make a model combining a few of those all together, and have all of these as predictors?? Like have "Gene" or "generaltumortype" as our response variable, and have the exposures/therapies ("smoke" and "ind_target_therapy" etc) together to see how well our model's results confirms their findings.


- which genes (Gene) predict specific types of tumors (generaltumortype)?


##### GENE AS OUTCOME ("y")

- which genes (Gene) have more mutations in smokers (smoke)?

- which genes (Gene) have more mutations in patients that have undergone cancer therapy (ind_target_therapy)(ind_immune_therapy)(XRT)(ind_cytotoxic_therapy)?

GOAL: see if our model predicts ASXL1 when there are smokers, and the other genes when there are cancer therapies?



##### QUESTIONS

This is more *classification*, not *regression*, and I think all we've done in class so far is regression (predicting home prices...). 



Models?  
- decision tree  
- lin reg  


When we've used decision trees in class/lab, it's still been in the context of regression, where our outcome variable is a continuous number. But it would make sense if we could use decision trees for classifcation, too...  
- we did this in class today?




##### OTHER

- add info from Gene Ontology, get information about certain classes of genes (all involved in the same processes / pathways, etc.)



##### prof. arnold's thoughts

- try to find some analysis from the paper to do
- easiest: add prediction to 1st project
- other models, other things to delve into


- do the stuff we did in class today  

- if it's not regression, how do we make that work?

- binary does/doesn't have mutation
  - mult features: smoker, radiation therapy, 
  - inference task in a classification setting. inference about a logistic regression
  - model will give conf intervals if we ask it. see if they include 0.
  - predictionns: of all the things we know, what mutations will there be
  - @ timepoints: 1st timepoint + other data + age + gender + smoker + timeframe in between + therapy.
    - not necessarily include that mutaion in gene A = = > gene B, not causes each other.
  - somewhat obvious extensions.
  
  
  - how it would be collected, what it means, totally realistic.
  - coding / modeling -- doesn't take that much time in comparison to the scope.
  -how to robustly tell teh story of what you figured out.
  - defend why.
  - here's the evidence why i think this is right. based on (1) original code, it must meean this or (2) according to underlying biology, this must be the relationship.
  - a GOOD PROJECT has robust narrative & a small amount of targetted, simple code. nothing requires a LOT of code, just the RIGHT code.
  - if you can go deep on this, figure out how to bound the risk (like figuring out what data made a figure), go for it.
  - GET RMD running, delete stuff until it doesn't work. help to identity what's needed for making figure X. 
  - exploring the code, you don't need a lisence if you're sharing the results. we're fine using it as is.

- geom_freq_poly()

- inference: which one is more important. how does intervention play out? Why would changing oxidation be different.
  - for people w/ autism, how doese % oxidation diffeer?
  - is this a *real* difference? or just a statistical artifact? 
  #### theoery based approach
  - data %>% specify(`% oxidized` ~ Group) %>% hypothesis(null = "independence") %>% generate(1000, type = "permute") %>% calculate("diff in means") %>% visualize()-- empirical difference in means
  ^^ if there's no difference, we would expect this range. Simulation-Based Null Distribution
  - to mark where we are: point estimate! see where we are. infer package. get_p_value -- null_distribution, ovs_state = oint_estimate, direction = "two-sided"
  - yet he's sceptical of this type of *analysis*: is this causal? idk? tempting to say that [x] causes [y]. 
  - expiremental design: if we did something that changed x, we can say that it was casual on y.
  - OR, an all else being equal situation. Matched case-control. Econometrics: instrumental variables to measure that even if you can't match up people. BUT, you need to have confidence that you're measuring everything, that you're measuring any hidden confounding variable.


- prediction: classify as autism or not. look @ patterns in coefficients. Queestion isn't what doees % oxidation of metabolites mean. Just trying to make good predictions.



